---
title: "Project 1"
date: "2023-06-25"
author: "Spencer Williams"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages Needed
```{r packages, include=TRUE}
library(httr)
library(dplyr)
library(jsonlite)
library(rmarkdown)
library(tidyverse)
```

### Installing Stock Data from January 9
```{r intro, include=TRUE}
library(dplyr)
myData <- GET("https://api.polygon.io/v2/aggs/grouped/locale/us/market/stocks/2023-01-09?adjusted=true&apiKey=BnFt9sqaLWehzloLkpMtEeOG4WjAqkUc")
str(myData, max.level = 1)
parsedJan <- fromJSON(rawToChar(myData$content))
str(parsedJan, max.level = 1)
parsedJan$results %>% colnames()
```

I am getting my data from the API using my key. I then created a data table with the results section from the API.

### Making A Data Frame for January 9th Data
```{r dataframe, include=TRUE}
parsedJanDataFrame <- as.data.frame(parsedJan)
colnames(parsedJanDataFrame) <- c("Query", "Total_Request", "Adjusted", "Symbol", "Volume", "Volume_Weight", "Open", "Close", "High", "Low", "Timestamp", "Transactions", "Status", "Request_ID", "Count")
```

I created a data frame for the January 9th data from the API, and relabeled the variable names.

### Installing Stock Data from March 9
```{r intro2, include=TRUE}
library(dplyr)
myData2 <- GET("https://api.polygon.io/v2/aggs/grouped/locale/us/market/stocks/2023-03-09?adjusted=true&apiKey=BnFt9sqaLWehzloLkpMtEeOG4WjAqkUc")
str(myData2, max.level = 1)
parsedMar <- fromJSON(rawToChar(myData2$content))
str(parsedMar, max.level = 1)
parsedMar$results %>% colnames()
```

I am getting my data from the API using my key. I then created a data table with the results section from the API.

### Making A Data Frame for March 9th Data
```{r dataframe2, include=TRUE}
parsedMarDataFrame <- as.data.frame(parsedMar)
colnames(parsedMarDataFrame) <- c("Query", "Total_Request", "Adjusted", "Symbol", "Volume", "Volume_Weight", "Open", "Close", "High", "Low", "Timestamp", "Transactions", "Status", "Request_ID", "Count")
```

I created a data frame for the March 9th data from the API, and relabeled the variable names.

### Combining the Two Data Frames and Adding a Month, Difference, and Return Variables
```{r combine, include=FALSE}
library(tidyr)
parsedJanDataFrame %>% drop_na()
parsedJanDataFrame <- mutate(parsedJanDataFrame, Month = "January", Difference = (Open - Close))
parsedMarDataFrame %>% drop_na()
parsedMarDataFrame <- mutate(parsedMarDataFrame, Month = "March", Difference = (Open - Close))
parsedFullData <- full_join(parsedJanDataFrame, parsedMarDataFrame)
```

I hid these statements because the list of data was very long. I dropped the na's from the dataset and created two new variables, month and difference between the opening stock price and closing stock price. I then did a full join of the two data frames from each month to create one big data frame.

```{r return, include=TRUE}
Returns <- vector()
for (i in seq_len(nrow(parsedFullData))) {
  if(parsedFullData$Difference[i] >= 0) {
    Returns[i] <- "Positive"
  } else if (parsedFullData$Difference[i] <= 0) {
    Returns[i] <- "Negative"
  } else {
    Returns[i] <- "Error"
  }
}
parsedFullData$Returns <- Returns

head(parsedFullData)
```

I created an else if statement that will determine whether the stocks in the full dataset had a positive outcome during the day.

### Creating a Contingency Table between Returns and Month
```{r contingency, include=TRUE}
table(parsedFullData$Returns, parsedFullData$Month)
```

This contingency table shows that overall, stocks had more of a positive growth on March 9th than January 9th. Although, both days shows positive growth, 85.6% of March 9th stocks increased compared to only 60.9% on January 9th.

### Creating a Contingency Table bewteen Month and Opening Price ($20.00)
```{r contingency2, include=TRUE}
Twenty <- vector()
for (i in seq_len(nrow(parsedFullData))) {
  if(parsedFullData$Open[i] >= 20) {
    Twenty[i] <- "Higher"
  } else if (parsedFullData$Open[i] <= 20) {
    Twenty[i] <- "Lower"
  } else {
    Twenty[i] <- "Error"
  }
}
parsedFullData$Twenty <- Twenty

table(parsedFullData$Twenty, parsedFullData$Month)
```

Based on this contingency table comparing the opening price of each stock to the month variable, a $20.00 stock seems to be around the median. In January, 5315 stocks were above an opening price of $20.00 while 5638 stocks were below this price. In March, these totals were 5417 above and 5427 below the opening price of $20.00.

### Creating Numerical Summaries Grouped by Month
```{r summary, include=TRUE}
tapply(parsedFullData$Low, parsedFullData$Month, summary)
tapply(parsedFullData$High, parsedFullData$Month, summary)
```

Here, I am comparing numerical summaries between months for the daily high and lows. Based on the mean and median of each month, the data seems to be heavily skewed to the right (higher stock price). The summary statistics are relatively similar for the two months at daily low and high prices.



